Off the beaten path: \\ Gaia reveals GD-1 stars outside of the main stream

### 2018-05-01 paper/gd1.tex
abstract edit
fig1 annotations

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold
streams of stars that are extremely valuable tracers of the large- and
small-scale distribution of mass in the Galaxy.
Using the exquisite data from the \gaia\ second data release combined with
\pans\ photometry, we present a sample of highly-probable members of the
longest cold stream in the Milky Way -- GD-1.
The resulting map of GD-1: (1) extends the apparent length of the stream by
$20^\circ$, (2) reveals a plausible location for the progenitor, (3) detects 
significant density variations along the stream, and (4) indicates
the existence of stream members off the main stream track.
The additional length and proper treatment of the progenitor will aid in dynamical modeling of GD-1 and map the Galaxy on large scales, while its complex morphology points to a turbulent history, potentially informing on massive, small-scale perturbers within 30\,kpc.

% Most stellar streams discovered around the Milky Way reside in the Galactic halo
% and are therefore primarily sensitive to the dark matter distribution.
% The longest and most prominent thin stream in the Milky Way --- the GD-1 stream --- has been used
% to constrain the shape of the inner halo, and is a promising place to search for
% signatures of dark matter subhalo interactions.
% selected to have
% small parallaxes and retrograde proper motions using data from the \gaia\
% mission, and \pans\ photometry consistent with an old and metal-poor population
% at a distance of $\sim10~\kpc$.
% This selection extends the stream by $20^\circ$ in apparent length, reveals a
% plausible location of the progenitor, and shows significant density variations
% along the stream in unprecedented detail.
% This new view of the stream reveals significant density variations
% along the stream in unprecedented detail.
% In addition to confirming several prominent gaps detected in photometric maps of
% GD-1, for the first time we find evidence of stream members off of the main
% track.
These features could be an indication of complex internal kinematics within the
progenitor unlike any of the known globular clusters, or a result of a stream
interaction with a massive perturber.
% Detailed modeling of the positional and velocity offsets
% or an indication of complex internal kinematics within the
% stream progenitor.
% Velocity offset between these features and the stream is an order of magnitude larger than the variance in velocity fields of known globular cluster, thus favoring the encounter origin.
% If this is confirmed, off-stream features will enable a robust estimate of the perturber's mass.
% If caused by a subhalo interaction, we estimate that the detected underdensity
% and the associated spur could have been induced by a perturbation of a
% $\sim10^7\,\rm M_\odot$(?) object $\sim 0.5\,$Gyr ago.
\end{abstract}

@@

\acknowledgements{
It is a pleasure to thank
Lauren Anderson,
Vasily Belokurov,
Andrew R. Casey,
Marla Geha,
David W. Hogg,
Benjamin D. Johnson,
Kathryn V. Johnston,
Sergey Koposov,
Mariangela Lisanti,
Edward Schlafly,
and David N. Spergel. % for useful discussions and feedback.

>>>>

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold
streams of stars that are extremely valuable tracers of the large- and
small-scale distribution of mass in the Galaxy.
Using the exquisite data from the \gaia\ second data release combined with
\pans\ photometry, we present a sample of highly-probable members of the
longest cold stream in the Milky Way -- GD-1.
The resulting map of GD-1: (1) extends the apparent length of the stream by
$20^\circ$, (2) reveals a plausible location for the progenitor, (3) detects 
significant density variations along the stream, and (4) indicates
the existence of stream members perturbed off the main stream track.
\gaia\
The additional length and proper treatment of the progenitor will aid in dynamical modeling of GD-1 and map the Galaxy on large scales.
The complex morphology of the stream points to a turbulent history; the detailed morphology in phase space of the perturbed stream members could potentially provide constraints on dark matter substructure that might have interacted with the stream.

% Most stellar streams discovered around the Milky Way reside in the Galactic halo
% and are therefore primarily sensitive to the dark matter distribution.
% The longest and most prominent thin stream in the Milky Way --- the GD-1 stream --- has been used
% to constrain the shape of the inner halo, and is a promising place to search for
% signatures of dark matter subhalo interactions.
% selected to have
% small parallaxes and retrograde proper motions using data from the \gaia\
% mission, and \pans\ photometry consistent with an old and metal-poor population
% at a distance of $\sim10~\kpc$.
% This selection extends the stream by $20^\circ$ in apparent length, reveals a
% plausible location of the progenitor, and shows significant density variations
% along the stream in unprecedented detail.
% This new view of the stream reveals significant density variations
% along the stream in unprecedented detail.
% In addition to confirming several prominent gaps detected in photometric maps of
% GD-1, for the first time we find evidence of stream members off of the main
% track.
% These features could be an indication of complex internal kinematics within the
% progenitor unlike any of the known globular clusters, or a result of a stream
% interaction with a massive perturber.
% Detailed modeling of the positional and velocity offsets
% or an indication of complex internal kinematics within the
% stream progenitor.
% Velocity offset between these features and the stream is an order of magnitude larger than the variance in velocity fields of known globular cluster, thus favoring the encounter origin.
% If this is confirmed, off-stream features will enable a robust estimate of the perturber's mass.
% If caused by a subhalo interaction, we estimate that the detected underdensity
% and the associated spur could have been induced by a perturbation of a
% $\sim10^7\,\rm M_\odot$(?) object $\sim 0.5\,$Gyr ago.
\end{abstract}

@@

\acknowledgements{
It is a pleasure to thank
Marla Geha,
David W. Hogg,
Kathryn V. Johnston;
Lauren Anderson,
Vasily Belokurov,
Andrew R. Casey,
Benjamin D. Johnson,
Sergey Koposov,
Mariangela Lisanti,
Edward Schlafly,
and David N. Spergel. % for useful discussions and feedback.
